---
title: "Black Carbon (BC) and PM2.5 Analysis"
author: "Wenlu Ye"
date: "9/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, fig.width = 12, fig.height = 4)
```

This analysis examines the personal exposure to PM2.5 and black carbon (BC) in three countries: Kenya, Ghana, and Bangladesh. The dataset contains PM2.5 and BC exposure samples collected at three time points: baseline, followup 1 and followup 2. An intervention was deployed to half of the participants after baseline visit. The current Markdown focuses on the descriptive analyses for PM2.5 and BC personal exposure, the correlation between PM2.5 and BC exposure as well as plots to understand the distribution of PM2.5 and BC personal exposure by country and study phase.

```{r load-data}
#setwd()
library(tidyverse)
library(data.table)
library(lubridate)
library(knitr)
pm_bc = fread("data.csv")
```

1. PM Concentration Descriptive Analysis
```{r PM descriptive}
#Remove negative values
pm_bc = subset(pm_bc, pm_conc_ug_m3 >= 0)
#Descriptive summary
summary_pm <- pm_bc %>% 
  group_by(country, phase) %>%
  summarise(N = n(), Min = min(pm_conc_ug_m3), Median = median(pm_conc_ug_m3), Mean = mean(pm_conc_ug_m3), SD = sd(pm_conc_ug_m3), Max = max(pm_conc_ug_m3))
knitr::kable(summary_pm, digits = 2)
```

2. BC Concentration Descriptive Analysis
```{r BC descriptive}
#Remove negative values
pm_bc = subset(pm_bc, bc_conc_ug_m3 >= 0)
#Descriptive summary
summary_bc <- pm_bc %>% 
  group_by(country, phase) %>%
  summarise(N = n(), Min = min(bc_conc_ug_m3), Median = median(bc_conc_ug_m3), Mean = mean(bc_conc_ug_m3), SD = sd(bc_conc_ug_m3), Max = max(bc_conc_ug_m3))
knitr::kable(summary_bc, digits = 2)
```

3. PM and BC Correlation 
```{r correlation}
p = ggplot(pm_bc, aes(x = bc_conc_ug_m3, y = pm_conc_ug_m3, color = phase)) +
  geom_point(alpha = 0.5, aes(size = phase)) + 
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE) +
  scale_shape_manual(values=c(3, 16, 17)) + 
  scale_color_manual(values=c('#999999','#E69F00', '#56B4E9')) +
  scale_size_manual(values=c(2,2,2)) +
  theme(legend.position="right") + 
  labs(x = expression(BC~personal~exposure~(µg/m^3)), y = expression(Filter~corrected~PM[2.5]~personal~exposure~(µg/m^3))) 
plot1 = p + facet_grid(cols = vars(country))
plot1

summary_corr <- pm_bc %>%
  group_by(country, phase) %>%
  summarise(Correlation = cor(bc_conc_ug_m3, pm_conc_ug_m3, method = 'spearman'))
knitr::kable(summary_corr, digits = 2)
```

4. PM Density Plot by country and study phase 
```{r density-pm}
p = ggplot(pm_bc, aes(pm_conc_ug_m3, fill = phase, color = phase)) +
  geom_density(alpha= .5) + 
  xlim(0, 600) + 
  scale_color_brewer(palette="Dark2") +
  scale_fill_brewer(palette="Dark2") +
  theme(legend.position="right") + 
  labs(x = expression(Filter~corrected~PM[2.5]~personal~exposure~(µg/m^3)), y = "Density") +
  geom_rug()
  
plot2 = p + facet_grid(cols = vars(country))
plot2
```

5. BC Density Plot by country and study phase 
```{r density-bc}
p = ggplot(pm_bc, aes(bc_conc_ug_m3, fill = phase, color = phase)) +
  geom_density(alpha= .5) + 
  xlim(0, 40) + 
  scale_fill_manual(values = c('#999999','#E69F00', '#56B4E9')) +
  scale_color_manual(values = c('#999999','#E69F00', '#56B4E9')) +
  theme(legend.position="right") + 
  labs(x = expression(BC~personal~exposure~(µg/m^3)), y = "Density") +
  geom_rug()
  
plot3 = p + facet_grid(cols = vars(country))
plot3
```


